### 2014-11-12 docs/ms.Rmd
Updated the ms and re-wrote parts of the crappy paragraphs that I wrote. Also added in some important notes about annominity in the data....which reminds me we need to remove these details from the data.

## Introduction
Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration. For these biologists, identifying best practice in measuring color and color pattern and in determining how other species see colour has been a great challenge. Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography and visual modeling.

The application of spectrophotometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966]. Before the widespread use of spectrophotometry, researchers assigned categorical color descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although no doubt applied carefully, these techniques limited progress in the field. For example, many researchers designed color assessment methods specifically for their projects [e.g. @aldrich_ecogeographic_1991] rendering data by different researchers incomparable. This was/is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX). Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994](MORE). Human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

Measuring reflectance spectra is now a reasonably inexpensive and convenient technique. Spectrophotometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field. However one drawback to spectrophotometry, is the ease with which erroneous data can be captured if attention is not paid to finer details of spectrum capture. For example, the angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008], can all contribute to potentially skewed data or alter results. Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject and to ensure repeatability, conditions under which data were collected must be reported exactly.

In conjunction with spectrophotometry, digital photography is increasingly being implemented in color quantification. High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images [@mckay_use_2013]. The main challenges for gathering and using color data from photographs are ensure correct standardization methods such as linearization are consistent lighting [@stevens_using_2007]. Crucially, most digital cameras used in color science are consumer products designed for human visual color space and hardware modifications for ultraviolet and infrared imaging are necessary [e.g. @chiao_visualization_2009; @mckay_use_2013], making photography complementary, rather than superior to spectrophotometry.

@@

Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013]. The potential for detrimental subjectivity in color measurement and the mathematical opacity of visual models for the average colour scienctist means that, as in any field, is it critical that methods are reported rigorously. Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011](Ask Rob Lanfear). The aims of this paper are to (a) review methods reporting in papers published on colour in 2013 from 20 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods. Our hope is that this instructional provides a useful set of guidelines to improve the reproducibility and transparency of studies on color.

@@

We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006] (Table 1 and 2). These included quantitative methods including spectrophotometric and photographic techniques, but not qualitative methods such as the use of Munsell chips or paint swatches. For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).

@@

To determine the current state of methods reporting in the field we searched for papers from 2013 in 23 top color science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Evology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_, _Ethology_). For each of the journals, we used the Boolean search terms 'colour*', 'color*' or 'spectra*' and searched for these in the title or abstract from each of the journal home pages. We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals. We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern; and 4) papers taking microspectrophotometric measurments of retinal absorbance. To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our scoring, papers were read and reassessed by a total of three authors. Any discrepancies between the assessor scores were discussed and agreed upon prior to analysis. We then tabulated what details of the methods were reported against our criteria.

@@

Our review of recent colour literarure highlights some major reporting gaps that will hinder the reproducibility of studies quantifying colour signals. While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly beleive that studies can do a much better job reporting details which are essential to the transparency and reproducibility of their work. When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investiagted expliclty stated. While all criteria are important some deserve special attanetion. In particular we found that colour scientists could do a much better job reporting light sources, number of pixel and spectra averaged used for analysis and the type of standards used. More improtantly, the distance and angle should be standard reporting for all studies using spectrophotomtry. These all have important consequences on downstream analysis and the reproducibility of colour signal quantification. As we all know, a picture can say 1000 words and we recommend the incoperation of spectral reflectance curves were possible as it provides insight on the colour signal, which is visually appealing and easy to understand by most readers. Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had an explicit figure in their paper showing these curves. Visual models were regularly used to analyse spectral data, something we welcome. However, there were many important aspects of the analysis that were seldom reported, inclding the type of quantum catch used, whether photoreceptor adaptation was accounted for and the type of neural noise used in receptor noise models. Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities.

In conclusion, our objective in this paper was not to point to single studies and lay blame, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons. Indeed, even in our own work [@umbers_bright_2013] we have failed to report important details, which in resptrospec can lead to confusion about how such data were collected and analysed. Nontheless, our critisim of reporting is meant to be constructive and is not to say that studies lacking improtant details are not excellent works, but it is only through critical appraisal and colleageal discussion that we can identify these gaps and constructively contribute to the growth of our field. Such assessments are particlularly important in an age where accountability, transparency and reproducibility of are utmost importance. We hope our critical evaluation of recent litearture has highlighted a need for more careful reporting of how colour signals have been measured and analysed. We beleive that part of the issue is simply that researchers seldom have clear guidelines to follow when undertaking research and miss what might be important aspects of their methods and analysis that need to be reported. We therefore hope that our paper will help estbalish more uniform reporting guidelines in the literature so that we can continue an exciting era of colour signal research in the life sciences.

>>>>

## Introduction
Since Thayer [-@thayer_concealing_1909], Poulton [-@poulton_colours_1890] and Wallace [-@wallace_natural_1891] generations of biologists have endeavored to explain the mechanisms and functions of animal and plant coloration. For these biologists, identifying best practices in measuring color and color pattern and in determining how other species see colour has been a great challenge. Over the past 150 years biologists, both independently and in collaboration with physicists and chemists, have developed a range of techniques used to measure and analyse animal and plant coloration including spectrophotometry (also called spectroradiometry, spectroscopy ETC), photography and visual modeling.

The application of spectrophotometry to questions in ecology and evolution revolutionized the field [@dyck_determination_1966]. Before the widespread use of spectrophotometry, researchers assigned categorical color descriptions, arbitrary rankings or scores [e.g. @aldrich_ecogeographic_1991; @wheelwright_colors_1985] or matched paint swatches or munsell chip ratings based on their own visual assessment [e.g. @burtt_analysis_1986; @key_temperature_1954]. Although when applied carefully they maybe useful, these techniques in general have limited progress in the field. For example, many researchers designed color assessment methods specifically for their projects [e.g. @aldrich_ecogeographic_1991] rendering data by different researchers incomparable. This was/is especially problematic when different measurement techniques are used on the same species (e,g, refXXX vs ref XXX). Also, human vision, insensitive to ultraviolet wavelengths, is inherently different to most other animals such as avian and reptilian tetrachromats, and insect trichromats [@bennett_sexual_1994](MORE). Human perception of color, and variation therein, is not only biologically irrelevant to most systems, but it can lead to erroneous conclusions [@bennett_sexual_1994; @endler_measurement_1990].

Measuring reflectance spectra is now a reasonably inexpensive and convenient technique fo quantifying color signals. Spectrophotometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers in the field. However one drawback to spectrophotometry, is the ease with which erroneous data can be captured if attention is not paid to finer details of spectrum capture. For example, the angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and object, the expectations of the researcher and the size, shape, glossiness or iridescence of the specimen [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008], can all contribute to potentially skewed data or alter results. Experience, guidance and objectivity are necessary to ensure data are consistent and representative of the subject. In addition, to ensure repeatability conditions under which data were collected must be reported exactly.

In conjunction with spectrophotometry, digital photography is increasingly being implemented in color quantification. High-resolution cameras are now inexpensive and portable, allow instantaneous sampling of multiple colour patches, and can produce highly repeatable images [@mckay_use_2013]. The main challenges for gathering and using color data from photographs are ensuring correct standardization methods, such as linearization and consistent lighting [@stevens_using_2007]. Crucially, most digital cameras used in color science are consumer products designed for human visual color space and hardware modifications for ultraviolet and infrared imaging are necessary [e.g. @chiao_visualization_2009; @mckay_use_2013], making photography complementary, rather than superior to spectrophotometry.

@@

Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013]. The potential for detrimental subjectivity in color measurement and the mathematical opacity of visual models for the average colour scienctist means that, as in any field, is it critical that methods are reported rigorously. Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011](Ask Rob Lanfear). The aims of this paper are to (a) provide a set of criteria for best-practices in reporting of the most common methods used to collect and analyse data and (b) assess how well we are doing in reporting these criteria. We review methods reporting in papers published on colour in 2013 from 23 top tier journals. Our hope is that this instructional provides a useful set of guidelines that colour scientists can follow that will improve the reproducibility and transparency of their work.

@@

We developed a list of metrics we deemed critical to be reported for reproducibility of color methods and analysis. Our guidelines are loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006] (Table 1 and 2). They included quantitative methods, such as spectrophotometric and photographic techniques, but not qualitative methods such as the use of Munsell chips or paint swatches. For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).

@@

To determine the current state of methods reporting in the field we searched for papers from 2013 in 23 top color science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Evology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_, _Ethology_). For each of the journals, we used the Boolean search terms 'colour*', 'color*' or 'spectra*' and searched for these in the title or abstract from each of the journal home pages. We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals. We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern; and 4) papers taking microspectrophotometric measurments of retinal absorbance. To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our scoring, papers were read and reassessed by a total of three authors. Any discrepancies between the assessor scores were discussed and agreed upon prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep. We have keep the papers used in our dataset annoymous to avoid accusing specific colleagues of inadvertant reporting. We feel that the issue of reporting is something attributable to the ambiguity and lack of guidelines in our field and not something specific to individual research groups.

@@

Our review of recent colour literarure highlights some important reporting gaps that will hinder the reproducibility of studies quantifying colour signals. While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly beleive that studies can do a much better job reporting details which are essential to the transparency and reproducibility of their work.  While all criteria are important some deserve special attanetion. In particular, we found that colour scientists could do a much better job reporting light sources, the number of pixels and spectra averaged and the type of standards used. More improtantly, the distance and angle should be standard reporting for all studies using spectrophotomtry. Visual models were regularly used to analyse spectral data, something we welcome. However, there were many important aspects of the analysis that were seldom reported, inclding the type of quantum catch used, whether photoreceptor adaptation was accounted for and the type of neural noise used in receptor noise models. Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities. These all have important consequences on downstream analysis and the reproducibility of colour signal quantification. While these criteria are important additional information can also be extremely helpful to readers. As we all know, a picture in worth 1000 words and we recommend the incoperation of spectral reflectance curves were possible as it provides insight on the colour signal, which is visually appealing and easy to understand by most readers. Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had an explicit figure in their paper showing these curves. When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investiagted expliclty stated.

In conclusion, our objective in this paper was not to lay blame on single studies, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross taxa comparisons. Indeed, even in our own work [e.g. @umbers_bright_2013] we have failed to report important details, which in resptrospec can lead to confusion about how such data were collected and analysed. Nontheless, our critisim of reporting is meant to be constructive and is not to say that studies lacking such details are not excellent works in their own right. There is no doubt many papers we evaluated have made important contributions. Nontheless, it is only through critical appraisal and colleageal discussion that we can identify these gaps and constructively contribute to the growth of our field. Such assessments are particlularly important in an age where accountability, transparency and reproducibility of are utmost importance. We hope our critical evaluation of recent litearture has highlighted a need for more thorough reporting of how colour signals have been measured and analysed. We beleive that part of the issue is simply that researchers seldom have clear guidelines to follow when undertaking research and as such miss important aspects of their methods and analysis that should to be reported. We therefore hope that our paper will help estbalish more uniform reporting guidelines in the literature so that we can continue an exciting era of colour signal research in the life sciences.
